Capitalized Title Here

### 2021-03-11 CMPSpaper_writing/CMPSpaper_writing.Rmd
added more citations in bib and revised "extract signature" section

In this paper we present an open-source implementation of the algorithm of the Congruent Matching Profile Segments (CMPS) method. @cmps developed the CMPS method for "objective comparison of striated tool marks" and demonstrated its use in some examples of comparing bullet signature correlations. Although @cmps described the CMPS algorithm  in their paper, conceptually, \hh{the authors did not release an implementation of their method}. Thus, our effort here is to introduce the CMPS method to the `R` community and provide an open-source implementation of the algorithm for people to use, review, and further improvements. Our implementation is wrapped up into  `R` package  `CMPS`, available on GitHub at \url{https://github.com/willju-wangqian/CMPS}.

@@

An important task of firearm examination is \hh{to answer the question of whether two pieces of evidence come from the same source or whether a piece of evidence matches a sample obtained from a specific firearm. Here, in particular, we are interested } to determine whether two bullets were fired from the same gun barrel. 
\hh{Assessing the similarity between two bullets is based on a comparison of  striation marks acquired during the firing process as bullets are propelled through the barrel. }
\hh{Current state of art sees firearms examiners make an assessment of similarity based on a visual comparison, generally, using a comparison microscope. This practice has been criticized for its lack of objectivity and the associated problem in calculating an error rate [@pcast].}
<!--It is achievable because lands in a gun barrel create striation marks on the bullet during the firing process, and it is believed that the same gun barrel will create highly similar striation marks on the bullets it fired. Typically the striation marks are created by lands in a barrel, so the striation marks are also sometimes refered as land engravings, and the areas the marks occupy are called land engraved areas (LEA). However, in recent years people became more cautious about firearms identification, and criticism of its subjective methodology arose.--> A report published by @nrc states that "[m]uch forensic evidence-including, for example, bite marks and firearm and toolmark identification—is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline." To overcome those criticisms and concerns, researchers have been making an effort to build databases [@nistdb] and develop frameworks and algorithms that  bring an objective and quantitative assessment into the field \hh{XXX cite some of these efforts - make sure to say that it is not an exhaustive list}. 
\hh{XXX The bullet signatures weren't mentioned before - make sure to include that there is a whole section that discusses how bullet signatures are extracted from scans.}
The bullet signatures extracted from the bullet are important in this process since they can well describe the striation marks on the bullet and can well be understood by algorithms.

In the following sections we will: review the background of bullet signature comparisons, discuss how we followed the idea described by @cmps for the implementation of the CMPS algorithm, and present results of applying our implementation to real data.

@@

The automated framework  used in this paper was proposed by @aoas. In order to obtain bullet signatures from the x3p files of bullet land impressions, we use the `R` packages `x3ptools` [@x3ptools] and `bulletxtrctr` [@bulletxtrctr]. `x3ptools` \hh{is a package to read, write, and generally, process } x3p files. 
<!--In the case of Hamby set 252, we need to convert the measurements of x3p files from meters to microns (`x3ptools::x3p_m_to_mum`) and rotate x3p files to make the long axis horizontal (`x3ptools::x3p_rotate`).--> 
\hh{it's land **engraved** areas, not impressions (that was wrong in Eric's first paper). Impressions are marks from impacts, such as the hit of a hammer or the hit from the breech face. Engravings are scratch marks.}
\autoref{fig:process}(a) shows a \hh{rendering of a 3d scan of a } bullet land  engraved area. The raised portion of the surface \hh{on the left and right of the scan are parts of the adjacent groove engraved areas (GEAs)}, \hh{the middle area shows a land-engraved area (LEA) with well expressed} striation marks. The first step of obtaining the bullet signature is to extract a cross-section at a fixed height along the land impression.
\hh{The next section is at an awkward level of detail - there's lots of details, but without the parameter settings it is still not a reproducible description. Could you instead describe things conceptually, and then link to the code that does this particular image?} `bulletxtrctr::x3p_crosscut_optimize` can be used to determine an ideal \hh{XXX ideal is a dangerous word} height, and `bulletxtrctr::x3p_crosscut` can be used to acquire the crosscut data (`ccdata`). \autoref{fig:process}(c) shows the typical pattern of the crosscut data, where "wedges" can be found on the left and right side and indicate the boundary between the land impression and the groove impression. `bulletxtrctr::cc_locate_grooves` is used to locate and chop these wedge areas so that only the curve of the land impression in the middle is kept. With the groove impressions excluded, we then use `bulletxtrctr::cc_get_signature` to smooth out the land impression curvature using LOESS (locally estimated scatter plot smoothing) [@loess] and obtain the bullet signature. \autoref{fig:process} illustrates this framework. Sometimes the algorithm in our framework cannot determine the ideal location of the crosscut or the grooves due to the deformation of a bullet, so we also developed a `R` Shiny App called `bulletinspectR` to identify and correct those errors. Note that NIST researchers used a different way of getting the bullet signatures in their paper [@cmps].

```{r bullet, echo=FALSE, out.width=".8\\textwidth", fig.cap="image of a XXX gun barrel and a fired bullet - \\hh{xxx my version of the image still has a groove engraved area marked as a land}"}
knitr::include_graphics("img/barrel_bullet.PNG", dpi = 100)
```

```{r process, echo=FALSE, out.width=".9\\textwidth", fig.cap="(a) front view of a scanned land engraved area (LEA). The optimized crosscut location is indicated by the white line. (b) view of the cross-section of the land engraved area at the white line in (a). (c) the crosscut data plotted in 2D; blue vertical lines indicate the position of left and right grooves (d) the crosscut data after chopping the left and right grooves (e) the fitted curvature using LOESS (f) after removing the curvature from the crosscut data, the bullet signature is obtained"}
knitr::include_graphics("img/figure1_v2.PNG", dpi = 100)
```

@@

The `CMPS` package contains a simple example to illustrate the basic usage of the package. The data in this example are twelve bullet signatures obtained from two bullets in Hamby set 252 [cite?] (Each bullet has six lands). The procedure for generating signatures from 3D bullet lands in x3p format is (algorithmic) [and has been discussed above]. And these two bullets are known to come from the same gun barrel, so for the 36 land-by-land comparisons, 6 of them are KM (known matching) comparisons and 30 are KNM (known non-matching) comparisons. To access the example data, one can use

@@

And anyone who is interested in the data source can find it in `bullets$source`. One can use these links to download the original x3p files from [NIST, cite?]. `bulletxtrctr::read_bullet(urllist = bullets$source)`

@@

In (NIST's original paper) that presents the CMPS algorithm researchers applied the CMPS method to one of the Hamby set scans (Hamby 252). In order to show that our implementation of the CMPS algorithm is able to reproduce the results in (original paper) and is able to be used for other data sets, we applied our implementation to both Hamby set 252 and Hamby set 44. Since the CMPS algorithm works with, in our case, bullet signatures, we conducted some data-processing. For both Hamby 252 and Hamby 44, we started with x3p files in the database, followed the framework proposed by Hare and Hofmann using the same set of parameters for both data sets, removed outliers, and obtained bullet signatures for each bullet land impression. Note that [in NIST's paper] they used a different framework to obtain bullet signatures. However, since NIST's work is not open-source, we were not able to follow their framework and were only able to reproduce the results for Hamby set 252 qualitatively.

>>>>

In this paper we present an open-source implementation of the algorithm of the Congruent Matching Profile Segments (CMPS) method. @cmps developed the CMPS method for "objective comparison of striated tool marks" and demonstrated its use in some examples of comparing bullet signature correlations. Although @cmps described the CMPS algorithm  in their paper, conceptually, \hh{the authors did not release an implementation of their method}. Thus, our effort here is to introduce the CMPS method to the `R` [@R] community and provide an open-source implementation of the algorithm for people to use, review, and further improvements. Our implementation is wrapped up into  `R` package  `CMPS`, available on GitHub at \url{https://github.com/willju-wangqian/CMPS}.

@@

An important task of firearm examination is \hh{to answer the question of whether two pieces of evidence come from the same source or whether a piece of evidence matches a sample obtained from a specific firearm. Here, in particular, we are interested } to determine whether two bullets were fired from the same gun barrel. 
\hh{Assessing the similarity between two bullets is based on a comparison of  striation marks acquired during the firing process as bullets are propelled through the barrel. }
\hh{Current state of art sees firearms examiners make an assessment of similarity based on a visual comparison, generally, using a comparison microscope. This practice has been criticized for its lack of objectivity and the associated problem in calculating an error rate [@pcast].}
<!--It is achievable because lands in a gun barrel create striation marks on the bullet during the firing process, and it is believed that the same gun barrel will create highly similar striation marks on the bullets it fired. Typically the striation marks are created by lands in a barrel, so the striation marks are also sometimes refered as land engravings, and the areas the marks occupy are called land engraved areas (LEA). However, in recent years people became more cautious about firearms identification, and criticism of its subjective methodology arose.--> A report published by @nrc states that "[m]uch forensic evidence-including, for example, bite marks and firearm and toolmark identification—is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline." To overcome those criticisms and concerns, researchers have been making an effort to build databases [@nistdb] and develop frameworks and algorithms that bring an objective and quantitative assessment into the field \hh{XXX cite some of these efforts - make sure to say that it is not an exhaustive list}. \hh{XXX The bullet signatures weren't mentioned before - make sure to include that there is a whole section that discusses how bullet signatures are extracted from scans.}
\wj{The notion of bullet signatures is one of the results of these efforts. Bullet signatures can be extracted from bullet land engraved areas (LEAs), typically one for each land engraved area, and how those bullet signatures are extracted from scans will be discussed in the Background section. Since bullet signatures can numerically describe the striation marks on the bullet and can be understood by algorithms, they played an important role in bringing objectivity into the field. }
In the following sections we will: review the background of bullet signature comparisons, discuss how we followed the idea described by @cmps for the implementation of the CMPS algorithm, and present results of applying our implementation to real data.

@@

The automated framework  used in this paper was proposed by @aoas. In order to obtain bullet signatures from the x3p files of bullet land engravings, we use the `R` packages `x3ptools` [@x3ptools] and `bulletxtrctr` [@bulletxtrctr]. `x3ptools` \hh{is a package to read, write, and generally, process } x3p files. 
<!--In the case of Hamby set 252, we need to convert the measurements of x3p files from meters to microns (`x3ptools::x3p_m_to_mum`) and rotate x3p files to make the long axis horizontal (`x3ptools::x3p_rotate`).--> 
\hh{it's land **engraved** areas, not impressions (that was wrong in Eric's first paper). Impressions are marks from impacts, such as the hit of a hammer or the hit from the breech face. Engravings are scratch marks.}
\autoref{fig:process}(a) shows a \hh{rendering of a 3d scan of a } bullet land  engraved area. The raised portion of the surface \hh{on the left and right of the scan are parts of the adjacent groove engraved areas (GEAs)}, \hh{the middle area shows a land-engraved area (LEA) with well expressed} striation marks. The first step of obtaining the bullet signature is to extract a cross-section at a fixed height along the land engraving.
\hh{The next section is at an awkward level of detail - there's lots of details, but without the parameter settings it is still not a reproducible description. Could you instead describe things conceptually, and then link to the code that does this particular image?}

\wj{The white horizontal line in \autoref{fig:process}(a) indicates this fixed height selected by an algorithm, and \autoref{fig:process}(b) shows the cross-sectional view of the bullet scan at this height. Then we can acquire the crosscut data (`ccdata`) of the scan at this height. \autoref{fig:process}(c) shows the typical pattern of the crosscut data, where "wedges" can be found on the left and right side and indicate the boundary between the land engraving and the groove engravings. The next step is to locate and chop these wedge areas so that only the data of the land engraving in the middle is kept (as shown in \autoref{fig:process}(d)). With the groove engravings excluded, we can then smooth out the land engraving curvature (\autoref{fig:process}(e)) using LOESS (locally estimated scatter plot smoothing) [@loess] and obtain the bullet signature (\autoref{fig:process}(f)). \autoref{fig:process} illustrates this framework. Sometimes the algorithm in our framework cannot determine the valid location of the crosscut or the grooves due to the deformation of a bullet, so we also developed a `R` Shiny App called `bulletinspectR` to identify and correct those errors. Note that NIST researchers used a different way of getting the bullet signatures in their paper [@cmps].}

<!-- `bulletxtrctr::x3p_crosscut_optimize` can be used to determine an ideal \hh{XXX ideal is a dangerous word} height, and `bulletxtrctr::x3p_crosscut` can be used to acquire the crosscut data (`ccdata`). \autoref{fig:process}(c) shows the typical pattern of the crosscut data, where "wedges" can be found on the left and right side and indicate the boundary between the land engraving and the groove engraving. `bulletxtrctr::cc_locate_grooves` is used to locate and chop these wedge areas so that only the curve of the land engraving in the middle is kept. With the groove engravings excluded, we then use `bulletxtrctr::cc_get_signature` to smooth out the land engraving curvature using LOESS (locally estimated scatter plot smoothing) [@loess] and obtain the bullet signature. \autoref{fig:process} illustrates this framework. Sometimes the algorithm in our framework cannot determine the ideal location of the crosscut or the grooves due to the deformation of a bullet, so we also developed a `R` Shiny App called `bulletinspectR` to identify and correct those errors. Note that NIST researchers used a different way of getting the bullet signatures in their paper [@cmps]. -->

```{r bullet, echo=FALSE, out.width=".8\\textwidth", fig.cap="image of a XXX gun barrel and a fired bullet - \\hh{xxx my version of the image still has a groove engraved area marked as a land}"}
knitr::include_graphics("img/barrel_bullet_ps.png", dpi = 100)
```

```{r process, echo=FALSE, out.width=".9\\textwidth", fig.cap="A framework of obtaining a bullet signature. (a) front view of a scanned land engraved area (LEA). The selected crosscut location is indicated by the white line. (b) view of the cross-section of the land engraved area at the white line in (a). (c) the crosscut data plotted in 2D; blue vertical lines indicate the position of left and right grooves. (d) the crosscut data after chopping the left and right grooves. (e) the fitted curvature using LOESS. (f) after removing the curvature from the crosscut data, the bullet signature is obtained"}
knitr::include_graphics("img/figure1_v2.PNG", dpi = 100)
```

@@

The `CMPS` package contains a simple example to illustrate the basic usage of the package. The data in this example are twelve bullet signatures obtained from two bullets in Hamby set 252 (Each bullet has six lands). The procedure for generating signatures from 3D bullet lands in x3p format is (algorithmic) [and has been discussed above]. And these two bullets are known to come from the same gun barrel, so for the 36 land-by-land comparisons, 6 of them are KM (known matching) comparisons and 30 are KNM (known non-matching) comparisons. To access the example data, one can use

@@

And anyone who is interested in the data source can find it in `bullets$source`. One can use these links to download the original x3p files from NIST Ballistics Toolmark Research Database [@nistdb]. `bulletxtrctr::read_bullet(urllist = bullets$source)`

@@

As presented in the work of @cmps, researchers applied the CMPS method to one of the Hamby set scans (Hamby 252). In order to show that our implementation of the CMPS algorithm is able to reproduce the results in (original paper) and is able to be used for other data sets, we applied our implementation to both Hamby set 252 and Hamby set 44. Since the CMPS algorithm works with, in our case, bullet signatures, we conducted some data-processing. For both Hamby 252 and Hamby 44, we started with x3p files in the database, followed the framework proposed by Hare and Hofmann using the same set of parameters for both data sets, removed outliers, and obtained bullet signatures for each bullet land engraving. Note that in the work of @cmps they used a different framework to obtain bullet signatures. However, since NIST's work is not open-source, we were not able to follow their framework and were only able to reproduce the results for Hamby set 252 qualitatively.
